Model Finding for Recursive Functions in SMT\thanks{ This research is partially supported by the Inria technological development action ``Contre-exemples Utilisables par Isabelle et Coq'' (CUIC). }

### 2015-05-07 tex/main.tex

\begin{abstract}
Model finding in SMT solvers is
generally limited to universally quantified formulas in some restricted fragment.
This paper introduces a translation which reduces axioms specifying a large
class of recursive functions, including well-founded (terminating),
to universally quantified formulas for which known
techniques are applicable.
An empirical evaluation confirms that the approach improves 
the performance of existing solvers on benchmarks from two sources. 
The translation is implemented as a preprocessor in the solver CVC4.
\end{abstract}

@@

Unfortunately, neither of these fragments can accommodate the vast majority of
quantified formulas that correspond to recursive function definitions: The
essentially uninterpreted fragment does not allow the argument of a
recursive function to be used inside a complex term on the right-hand side,
whereas the finite model finding techniques of Reynolds et al.\ are not
applicable for functions over infinite domains such as the integers or
algebraic datatypes. A simple example where both approaches fail is
\begin{equation}\label{eq:pii}
\forall n : {\typeint}.\;\, \lite\bigl(
n \leq 0,\allowbreak\; \con{p}( n ) \teq 1,\allowbreak\; \con{p}( n ) \teq 2 * \con{p}( n - 1 ) \bigr)
\end{equation}
This state of affairs is highly unsatisfactory, given the frequence of
recursive definitions in practice and the impending addition of a dedicated
command, \texttt{define-fun-rec}, to the SMT-LIB standard \cite{smtlib25}.

In this paper, we present a method for translating formulas involving recursice function
definitions into formulas where finite model finding techniques can be applied.
The recursive functions must meet a semantic criterion to be admissible
(Section~\ref{sec:prelim}). This criterion is met by well-founded (terminating)
recursive function definitions. It is not met by inconsistent
definitions such as $\forall n : {\typeint}.\;\, \con{f}(n) \teq \con{f}(n) +
1$. %ct and it is possible to construct consistent examples that do not meet it.
%as well as other examples of consistent definition.

>>>>

\begin{abstract}
SMT solvers have recently been extended with techniques for finding models
in presence of universally quantified formulas in some restricted fragments.
This paper introduces a translation which reduces axioms specifying a large
class of recursive functions, including well-founded (terminating),
to universally quantified formulas for which known
techniques are applicable.
An empirical evaluation confirms that the approach improves 
the performance of existing solvers on benchmarks from two sources. 
The translation is implemented as a preprocessor in the solver CVC4.
\end{abstract}

@@

Unfortunately, neither of these fragments can accommodate the vast majority of
quantified formulas that correspond to recursive function definitions: The
essentially uninterpreted fragment does not allow the argument of a
recursive function to be used inside a complex term on the right-hand side,
whereas the finite model finding techniques of Reynolds et al.\ are not
applicable for functions over infinite domains such as the integers or
algebraic datatypes. A simple example where both approaches fail is
\begin{equation}\label{eq:pii}
\forall x : {\typeint}.\;\, \lite\bigl(
x \leq 0,\allowbreak\; \con{p}( x ) \teq 1,\allowbreak\; \con{p}( x ) \teq 2 * \con{p}( x - 1 ) \bigr)
\end{equation}
This state of affairs is highly unsatisfactory, given the frequence of
recursive definitions in practice and the impending addition of a dedicated
command, \texttt{define-fun-rec}, to the SMT-LIB standard \cite{smtlib25}.

In this paper, we present a method for translating formulas involving recursice function
definitions into formulas where finite model finding techniques can be applied.
The recursive functions must meet a semantic criterion to be admissible
(Section~\ref{sec:prelim}). This criterion is met by well-founded (terminating)
recursive function definitions. It is not met by inconsistent
definitions such as $\forall x : {\typeint}.\;\, \con{f}(x) \teq \con{f}(x) +
1$. %ct and it is possible to construct consistent examples that do not meet it.
%as well as other example of consistent definition.
